# Developing an international network

```{r setup, include=FALSE}
WORDS_TO_IGNORE <- c("Analytics", "eRum", "hackseq", "Recomb", "Rencontres")
source("knitr-options.R")
source("spelling-check.R")
```

## Objective

There are many advantages of having a large international network. For example,

- I can get opportunities of work and collaboration,

- I can share my open source work to give back to the community,

- I get motivation from people recognizing my work.

I want to develop an international network in two different communities:

- that of programming language R (~~in priority~~) and
 
- that of the analysis of genetic data (my subject of thesis). 

In order to achieve this, I am trying to become more and more **visible on the international scene** in many different ways.

## Getting more visible

### Check this using Google Analytics and R

In order to get a sense of who is interested in me and my work, I decided to follow the flow of people on my websites using Google Analytics.

I can access Google Analytics API in R:

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(googleAnalyticsR)
ga_auth(".httr-oauth")
select(ga_account_list(), c(webPropertyName, websiteUrl, viewId)) 
```


### Website and blog

First, just before the beginning of my thesis, I created my personal website and my blog about R. I try to blog fairly regularly about R and my blog is relayed on R-bloggers, a platform that brings together ~750 R bloggers and is read by thousands of people internationally. 

More and more people are coming to by blog:

```{r}
data_blog <- google_analytics(
  viewId = "152155542",
  date_range = ymd("2016-10-01", Sys.Date()),
  dimensions = c("date"),  # , "pagePath", "hour", "medium"
  metrics = c("sessions")  # , "pageviews"
)

# No data from Google Analytics before this date
ga_begin <- which(data_blog$date == "2017-06-04")

qplot(y = cumsum(data_blog$sessions)) + 
  geom_vline(xintercept = ga_begin, color = "red") + 
  bigstatsr::theme_bigstatsr() + 
  labs(x = "Day of blogging", y = "Cumulative numbers of session on my blog")
```


### GitHub

GitHub is a platform where you can put your code. GitHub is a good way to make your work visible internationally and facilitate collaborations. On GitHub, I'm trying to create useful R packages and contribute to other packages. 

### Twitter, an important tool for work

I also try to be visible on Twitter, which is used by both communities that are of interest for me to communicate about new stuff.

To convince you, two tweets about my R course:

<blockquote class="twitter-tweet" data-lang="en" align="center"><p lang="en" dir="ltr">Teaching an advanced R course <a href="https://t.co/2pMG2FWcPs">https://t.co/2pMG2FWcPs</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/DataScience?src=hash&amp;ref_src=twsrc%5Etfw">#DataScience</a></p>&mdash; R-bloggers (@Rbloggers) <a href="https://twitter.com/Rbloggers/status/979539984679161857?ref_src=twsrc%5Etfw">30 mars 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


<blockquote class="twitter-tweet" data-lang="en" align="center"><p lang="en" dir="ltr">If you are at the airport today waiting for your ✈️, here&#39;s something good to read: <br>&quot;Advanced <a href="https://twitter.com/hashtag/RStats?src=hash&amp;ref_src=twsrc%5Etfw">#RStats</a> Course&quot; by <a href="https://twitter.com/privefl?ref_src=twsrc%5Etfw">@privefl</a>  <a href="https://t.co/JXKBI5lON2">https://t.co/JXKBI5lON2</a></p>&mdash; Colin Fay (@_ColinFay) <a href="https://twitter.com/_ColinFay/status/997073282905067520?ref_src=twsrc%5Etfw">17 mai 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

In parallel, let's plot the number of visitors of my course:

```{r}
data_course <- google_analytics(
  viewId = "172000044",
  date_range = ymd("2018-03-20", Sys.Date()),
  dimensions = c("date"),  # , "pagePath", "hour", "medium"
  metrics = c("sessions")  # , "pageviews"
)

ggplot(data_course) + 
  geom_vline(xintercept = ymd("2018-03-30", "2018-05-17"), 
             linetype = 2, color = "red") + 
  geom_point(aes(date, sessions)) + 
  bigstatsr::theme_bigstatsr()
```

Here, you can clearly see two peaks of visit of [my advanced R course](https://privefl.github.io/advr38book/), one when I blogged about it, which was relayed by [R-bloggers](https://twitter.com/Rbloggers) (59K followers), and the other peak when [Colin Fay](https://twitter.com/_colinfay) (4600+ followers) tweeted about it.

### Attending conferences

To develop an international network, it is also very important to attend conferences, which I'm also trying to do.

<blockquote class="twitter-tweet" data-lang="en" align="center"><p lang="en" dir="ltr">Use bigstatsr <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> package for matrices larger than 100GB. Great ⚡ talk by <a href="https://twitter.com/privefl?ref_src=twsrc%5Etfw">@privefl</a> at <a href="https://twitter.com/erum2018?ref_src=twsrc%5Etfw">@erum2018</a> <a href="https://twitter.com/hashtag/erum2018?src=hash&amp;ref_src=twsrc%5Etfw">#erum2018</a> <a href="https://twitter.com/hashtag/DataScience?src=hash&amp;ref_src=twsrc%5Etfw">#DataScience</a> <a href="https://t.co/magBeDTyVz">pic.twitter.com/magBeDTyVz</a></p>&mdash; Peter Laurinec (@petolauri) <a href="https://twitter.com/petolauri/status/996338652820201472?ref_src=twsrc%5Etfw">15 mai 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

This is me on stage in Budapest in May 2018. Before this conference, I had 118 followers on Twitter; after 3 days of conference, I had 158 (+34%).

Conferences I participated in:

- [Rencontres R 2018](https://r2018-rennes.sciencesconf.org/): The R package bigstatsr: Memory- and Computation-Efficient Statistical Tools for Big Matrices [[Slides]](https://privefl.github.io/RR18/bigstatsr.html)

- [eRum 2018](https://2018.erum.io/): An R package for statistical tools with big matrices stored on disk. [[Recording]](https://youtu.be/w3a1-KLmDM8) [[Slides]](https://privefl.github.io/eRum-2018/slides.html)

- [Recomb-Genetics 2018](http://recomb2018.fr/recomb-genetics/): Predicting complex diseases: performance and robustness. [[Slides]](https://privefl.github.io/thesis-docs/recomb18.html)

- [LIFE 2018](https://life.univ-grenoble-alpes.fr/life-world/events-news/life-annual-meeting-725690.htm?RH=11210192165071361): Predicting complex diseases: performance and robustness. [[Slides]](https://github.com/privefl/paper2-PRS/blob/master/LIFE/pres-f-priv%C3%A9.pdf)

- [hackseq 2017](http://www.hackseq.com/): Developing advanced R tutorials for genomic data analysis. [[Website]](https://hackseq.github.io/2017_project_5/)

- [useR!2017](https://user2017.brussels/): The R package bigstatsr:
Memory- and Computation-Efficient Tools for Big Matrices. [[Recording]](https://t.co/aYt0q8MeXJ)

### A local network

Finally, in addition to taking part in an international community, I think it's very important to take part in a more local community. That's one reason why I launched the initiative to make an R community in Grenoble, which started to be active in September 2017 (https://r-in-grenoble.github.io/).

## What's make people be interested in me?

Is this my blog posts? the conferences I attend? or the paper I wrote?

```{r}
data_about <- google_analytics(
  viewId = "153879696",
  date_range = ymd("2017-06-26", Sys.Date()),
  dimensions = c("date"),  # , "pagePath", "hour", "medium"
  metrics = c("sessions")  # , "pageviews"
)

blog_dates <- list.files("../blog/_posts/") %>%
  str_sub(end = 10) %>%
  parse_date() %>% 
  print()

conf_dates <- ymd("2017-07-04", "2017-07-07",  # useR!
                  "2018-04-19", "2018-04-20",  # recomb
                  "2018-05-14", "2018-05-16")  # eRum

papers_dates <- ymd("2017-09-17", "2018-05-05")

data_about %>%
  arrange(desc(sessions)) %>%
  head()
```

```{r}
important_dates <- data.frame(date = c(blog_dates[-(1:7)], conf_dates, papers_dates))

ggplot(data_about, aes(yday(date), sessions)) +
  facet_grid(format(date, format = "%Y") ~ .) + 
  geom_vline(aes(xintercept = yday(date)), data = important_dates,
             color = "blue", linetype = 2, size = 1) +
  geom_point(size = 2) +
  geom_line(aes(group = 1), size = 0.8) +
  geom_smooth(method = "loess", span = 0.25, color = "red") +
  bigstatsr::theme_bigstatsr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Day number", y = "Number of sessions on my #About webpage")
```

It seems that blogging about my advanced R course materials is what got me the largest number of people coming to my webpage to learn more about me. 

## Conclusion

Thus, my international network is under development.

I'm using my blog, Twitter, GitHub, Stack Overflow and conferences to get people to know me and want to work with me.

We can consider that it is a process that takes time, yet the indicators show that I am on the right track.


